Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-02-12 paper.tex
agregue unas correcciones de Dorso mas complicadas (se pueden modificar)

The complex behavior of pedestrians features
 either his (her) feelings and 
the environmental conditions. The former is expressed, for example, by his 
(her) moving ``attitude'' (say, self-assuredness). The latter brings out the 
observed separation between pedestrians. Also, the ``contacts'' between 
individuals \textcolor{blue}{produce} some kind of ``unwanted'' slowing down. All these 
observed patterns are commonly quantified in the literature into a set of 
characteristic  parameters. The experimental meaning of these parameters is as 
follows.

@@

\item[(ii)] Despite the reactive attitude $\tau$, the pedestrian \textcolor{blue}{adopts} a 
``free'' (undisturbed) walking speed $v_d$. This speed expresses his (her) 
motivation or intention to reach a certain destination (as comfortable as 
possible). Observations commonly associate $0.6\,$m/s, $1\,$m/s or $1.5\,$m/s 
to relaxed, normal or nervous walking speeds, respectively  
\cite{helbing_1995,helbing_2000,li_2015}.

@@

The maximum (realistic) compressions may be computed from the \textcolor{blue}{hook's} relation 
$F(x)=k_n(x)\,x$ and the compressibility $k_n(x)$ reported in Table~\ref{table_data}. 
An ``uncomfortable'' body force  $10\,$N $-$ $30\,$N can address compressions in the 
range of $0.030-0.055\,$m. Also, a ``hitting'' force of $60\,$N can address 
compressions between $0.045$ and $0.065\,$m. \\
%Thus, according to Table~\ref{table_data}, \textcolor{blue}{we may expect experimental values for $k_n(x)$ up to 
%$1000\,$kg/s$^2$ (for $F=30\,$N), or, up to $1400\,$kg/s$^2$ (for $F=60\,$N).} \\

@@

The equation of motion (\ref{eqn_motion}) can be written in terms of these 
(\textcolor{blue}{dimensionless}) magnitudes, while only three (reduced) parameters are needed.\\

>>>>

The complex behavior of pedestrians features
 either his (her) feelings and 
the environmental conditions. The former is expressed, for example, by his 
(her) moving ``attitude'' (say, self-assuredness). The latter brings out the 
observed separation between pedestrians. Also, the ``contacts'' between 
individuals \textcolor{blue}{produce} some kind of ``unwanted'' slowing down. All these 
observed patterns are commonly quantified in the literature into a set of 
characteristic  parameters. \textcolor{blue}{It is worth mentioning that the empirical
data reported in this Section corresponds to non-panicking situations in which pedestrians 
do not experience anxiety. }
The experimental meaning of these parameters is as follows.

@@

\item[(ii)] Despite the reactive attitude $\tau$, the pedestrian \textcolor{blue}{adopts} a 
``free'' (undisturbed) walking speed $v_d$. This speed expresses his (her) 
motivation or intention to reach a certain destination (as comfortable as 
possible). Observations commonly associate $0.6\,$m/s, $1\,$m/s or $1.5\,$m/s 
to relaxed, normal or nervous walking speeds, respectively  
\cite{helbing_1995,helbing_2000,li_2015}. 
\textcolor{blue}{Nevertheless, when pedestrians are in a panic situation,
 the actual speed tends to be lower than the desired velocity $v_d$. }

@@

The maximum (realistic) compressions may be computed from the \textcolor{blue}{Hook's} relation 
$F(x)=k_n(x)\,x$ and the compressibility $k_n(x)$ reported in Table~\ref{table_data}. 
An ``uncomfortable'' body force  $10\,$N $-$ $30\,$N can address compressions in the 
range of $0.030-0.055\,$m. Also, a ``hitting'' force of $60\,$N can address 
compressions between $0.045$ and $0.065\,$m. \\
%Thus, according to Table~\ref{table_data}, \textcolor{blue}{we may expect experimental values for $k_n(x)$ up to 
%$1000\,$kg/s$^2$ (for $F=30\,$N), or, up to $1400\,$kg/s$^2$ (for $F=60\,$N).} \\

@@

The equation of motion (\ref{eqn_motion}) can be \textcolor{blue}{rewritten} in terms of these 
(\textcolor{blue}{dimensionless}) magnitudes, while only three (reduced) parameters are needed.\\
